---
title: "assignment-7"
author: "Gisela Noguera Toffoletti"
output: github_document
---

Assignment 7: Relational data plus revisiting data wrangling and
visualization
================

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(dslabs)
```

<br>

## Excercise: 2016 election result and polling

For this exercise, we will explore the result of the 2016 US
presidential election as well as the polling data. We will use the
following three datasets in the `dslabs` package, and use `join`
function to connect them together. As a reminder, you can use `?` to
learn more about these datasets.

- `results_us_election_2016`: Election results (popular vote) and
  electoral college votes from the 2016 presidential election.

- `polls_us_election_2016`: Poll results from the 2016 presidential
  elections.

- `murders`: Gun murder data from FBI reports. It also contains the
  population of each state.

We will also use [this
dataset](https://raw.githubusercontent.com/kshaffer/election2016/master/2016ElectionResultsByState.csv)
to get the exact numbers of votes for question 3.

<br>

### Question 1. What is the relationship between the population size and the number of electoral votes each state has?

**1a.** Use a `join` function to combine the `murders` dataset, which
contains information on population size, and the `results_us_election_2016` dataset, which contains information on the
number of electoral votes. Name this new dataset `q_1a`, and show its
first 6 rows.

```{r}
q_1a <- murders |> 
  left_join(results_us_election_2016, join_by(state))
head(q_1a)
```

<br> <br>

**1b.** Add a new variable in the `q_1a` dataset to indicate which
candidate won in each state, and remove the columns `abb`, `region`, and `total`. Name this new dataset `q_1b`, and show its first 6 rows.

```{r}
q_1b <- q_1a |> 
  mutate(winner = ifelse(clinton > trump, "clinton", "trump")) |> 
  select(state, population, electoral_votes, clinton, trump, others, winner)
head(q_1b)
```

<br> <br>

**1c.** Using the `q_1b` dataset, plot the relationship between
population size and number of electoral votes. Use color to indicate who won the state. Fit a straight line to the data, set its color to black, size to 0.1, and turn off its confidence interval.

```{r, message=FALSE, warning=FALSE}
q_1b |> 
  ggplot(mapping = aes(x = population, y = electoral_votes, colour = winner)) +
  geom_point() +
  geom_smooth(method = "lm", se = F, linewidth = 0.1, colour = "black")
```

<br> <br>

### Question 2. Would the election result be any different if the number of electoral votes is exactly proportional to a state’s population size?

**2a.** First, convert the `q_1b` dataset to longer format such that the `population` and `electoral_votes` columns are turned into rows as shown below. Name this new dataset `q_2a`, and show its first 6 rows.

```{r}
q_2a <- q_1b |> 
  pivot_longer(c("population", "electoral_votes"), names_to = "metric", values_to = "value") |> 
  select(state, clinton, trump, others, winner, metric, value)
head(q_2a)
```

<br> <br>

**2b.** Then, sum up the number of electoral votes and population size
across all states for each candidate. Name this new dataset `q_2b`, and print it as shown below.

```{r, message=FALSE}
q_2b <- q_2a |> 
  group_by(metric, winner) |> 
  summarise(
     value = sum(value))
q_2b
```

<br> <br>

**2c.** Use the `q_2b` dataset to contruct a bar plot to show the final electoral vote share under the scenarios of **1)** each state has the number of electoral votes that it currently has, and **2)** each state has the number of electoral votes that is exactly proportional to its population size. Here, assume that for each state, the winner will take all its electoral votes.

<br>

*Hint: `geom_col(position = "fill")` might be helpful.*

```{r}
q_2b |> 
  ggplot(aes(metric, value, fill = winner)) +
  geom_col(position = "fill")
```


<br> <br>




